---
title: "P8131_hw3"
author: "Apoorva Srinivasan"
date: "2/25/2019"
output: pdf_document
---


```{r}
library(tidyverse)
```


##Problem 1


```{r}
case = tibble(
  age = c(25, 35, 45, 55, 65, 75),
  non_exposed = c(0, 5, 21, 34, 36, 8),
  exposed = c(1, 4, 25, 42, 19, 5)
) %>% 
  mutate(study = 'case')
control = tibble(
  age = c(25, 35, 45, 55, 65, 75),
  non_exposed = c(106, 164, 138, 139, 88, 31),
  exposed = c(9, 26, 29, 27, 18, 0)
) %>% 
  mutate(study = 'control')
```

##### Data manipulation


```{r}
p1 = 
  rbind(case, control) %>%
  gather(key = 'exposure_status', value = 'number', non_exposed:exposed) %>% 
  mutate(exposure_status = recode(exposure_status, non_exposed = '0', exposed = '1')) %>% 
  mutate(exposure_status = as.factor(exposure_status)) %>% 
  spread(key = study, value = number)  %>% 
  select(case, control, exposure_status, age)
p1
```

##### Model fitting

```{r}
fit = glm(cbind(case, control) ~ exposure_status + age, family = binomial(link = 'logit'), p1)
summary(fit)
```

#### Interpretation

* The log odds ratio of esophageal cancer for low versus high daily alcohol consumption is 1.78, holding age constant. 
* The log odds ratio of esophageal cancer for one-year increase in age is 0.06, holding exposure status constant. 

##### Hypothesis test

Model 0: (case, control) ~ age

Model 1: (case, control) ~ age + alcohol_exposure

```{r}
fit_0 = glm(cbind(case, control) ~ age, family = binomial(link = 'logit'), p1)
G0 = sum(residuals(fit, type = 'pearson')^2) # pearson chisq 
phi = G0 / 9
test_statistic = (fit_0$deviance - fit$deviance) / phi
test_statistic
1 - pf(test_statistic, 1, 9) 
```

The p-value for F test is 0.00036 < 0.05. Hence, we reject the null and conclude that Model 1 is the better model. 



## Problem 2


```{r}
  y_bean_oa75 = c(10, 23, 23, 26, 17)
  m_bean_oa75 = c(39, 62, 81, 51, 39)
  y_bean_oa73 = c(8, 10, 8, 23, 0)
  m_bean_oa73 = c(16, 30, 28, 45, 4)
 
  
bean = tibble(
    y = c(y_bean_oa75, y_bean_oa73),
    m = c(m_bean_oa75, m_bean_oa73),
    extract = 'bean',
    seed = c(rep('75', length(y_bean_oa75)), rep('73', length(y_bean_oa73)))
) 


y_cucbr_oa75 = c(5, 53, 55, 32, 46, 10)
m_cucbr_oa75 = c(6, 74, 72, 51, 79, 13)
y_cucbr_oa73 = c(3, 22, 15, 32, 3)
m_cucbr_oa73 = c(12, 41, 30, 51, 7)
  
cucumber = tibble(
    y = c(y_cucbr_oa75, y_cucbr_oa73),
    m = c(m_cucbr_oa75, m_cucbr_oa73),
    extract = 'cucumber',
    seed = c(rep('75', length(y_cucbr_oa75)), rep('73', length(y_cucbr_oa73))) 
    )
p2 = 
  rbind(bean, cucumber) %>% 
  mutate(x = m - y) %>% 
  select(y, x, seed, extract)
  
```

####1)

```{r}
fit_glm = glm(cbind(y, x) ~ seed + extract, family = binomial(link = 'logit'), p2)
summary(fit_glm)
```

##### Interpretation

* The intercept -0.7005 is the odds of germination for O. aegyptiaca 73 on bean root extract. 

* The odds ratio of germination for O. aegyptiaca 75 versus O. aegyptiaca 73 is 0.2705, holding the type of root extract constant.

* The odds ratio of germination for cucumber root extract versus bean root extract is 1.0647, holding the type of seed constant. 


####2) Overdispersion

```{r}
pval = 1 - pchisq(fit_glm$deviance, 18)
pval #bad fit
```

The result above(p value smaller than 0.05) indicates that the fit is not good. Plotting graph to check if over dispersion exists.

```{r }
res = residuals(fit_glm, type = 'pearson')
plot(qnorm((21 + 1:21 + 0.5)/(2*21 + 1.125)), sort(abs(res)),xlab = 'Expected Half-Normal Order Stats', ylab = 'Ordered Abs Pearson Residuals')
abline(a = 0, b = 1)
abline(a = 0, b = sqrt(phi), lty = 2)
```

We can see from the graph above that there is constant over-dispersion.

```{r }
G.stat = sum(residuals(fit_glm, type = 'pearson')^2) 
G.stat
phi = G.stat / 18
phi

summary(fit_glm, dispersion = phi)
```


####Interpretation:

Same interpretaion for the coefficients as mentioned above. The standard errors are enlarged hence, the p-values for the coefficients become larger. 

####3)


Overdispersion occurs when the underlying distributions are not independantly and identically distributed. In this case, it may be due to correlation of germination rate of seeds within each batch. The germination rates are dependent because the seeds within a batch share the same environment. Moreover, the heterogeneity of living conditions in different replicates may add to the variance in germination rate.
