---
title: "Bm2_hw5"
author: "Apoorva Srinivasan"
date: "3/6/2019"
output: pdf_document
---

```{r}
library(tidyverse)
library(MASS)
library(pscl)
library(nnet)
```

```{r}
crab_data = read.table("HW5-crab.txt", header = TRUE) %>%
  as.tibble() %>%
  mutate(C = as.factor(C),
         S = as.factor(S)
  )
```

```{r}
m1 <- glm(Sa~W, family=poisson(link = log), data=crab_data)
summary(m1)
```


the log rate of no of satellites increases by 0.164 with a unit inc in female's width of carapace

OR 

the number of satellites increases by 1.18 with a unit increase in female's width of carapace

```{r}
m2 = glm(Sa~W + Wt, family=poisson(link = log), data=crab_data)
summary(m2)
```



Adjusting for weight of the females, the number of satellites increases by 1.05 with a unit increase in width of carapace.
Adjusting for width of carapase, the number of satellites increases by 1.57 for a unit increase in weight of females.



####deviance analysis

```{r}
test.stat=m1$deviance-m2$deviance
df=171-170
pval=1-pchisq(test.stat,df=df) # chisq test
pval #rej, go with the bigger model
```

The p value is smaller than 0.05. Hence, we reject the null and go with the bigger model


####Overdispersion

```{r}
pval = 1 - pchisq(m2$deviance, df =170)
pval 
```

p-value is 0 indicating that it's a bad fit

```{r}
res.m2 = residuals(m2,type='pearson',data=crab_data)  
res.m2
G1=sum(res.m2^2) 
phi=G1/(170)
phi #3.16
m2$deviance/m2$df.residual #3.3
```


```{r}
plot(qnorm((173+1:173+0.5)/(2*173+1.125)),sort(abs(res.m2)),xlab='Expected Half-Normal Order Stats',ylab='Ordered Abs Pearson Residuals')
abline(a=0,b=1)
abline(a=0,b=sqrt(phi),lty=2) 
```


We can see from the graph above that over-dispersion occurs.

```{r}
m2_nb=glm.nb(Sa~W + Wt, data=crab_data)
summary(m2_nb)
```

####Interpretation:

the number  of satellites increases by 1.02 with a unit increase in female's width of carapace and the number of satellites increases by 1.93 with a unit increase in weight of the female crab

Since the dispersion parameter for Negative Binomial is very small(0.9324) it means that it's been adjusted for overdispersion

##PROBLEM 2

```{r}
parasite_data = read.table("HW5-parasite.txt", header = TRUE) %>%
  as.tibble() %>%
  janitor::clean_names() %>%
  dplyr::select(sample, intensity, year, length, area) %>%
  mutate(area = as.factor(area),
         year = as.factor(year)
  )
  
```

```{r}
p2a = glm(intensity~area + year + length , family=poisson(link = log), data=parasite_data)
summary(p2a)
```

####Interpretation:

* The log rate ratio of intensity between area1 and area2 is -0.212. This implies that the number of parasites in area 2 is 0.809 times area1

* The number of parasites in area 3 is estimated to be 0.890 times the number in area 1. 

* The number of parasites in area 4 is estimated to be 4.075 times the number in area 1.


* The number of parasites in year 2000 is 1.955 times the number in 1999, holding other variables constant. 

* The number of parasites in year 2001 is 0.804 times the number in 1999.

* Holding other variables constant, the number of parasites changes 0.972 fold with one-unit increase in length of the fish

```{r}
G = sum(residuals(p2a, type = 'pearson')^2)
1 - pchisq(G, df = p2a$df.residual)
```

The p value is <0.05 then we conclude that the fit is not good


```{r}
p2c = zeroinfl(intensity ~ area + year + length | area + year + length, data = parasite_data)
summary(p2c)
```



####Interpretation:

* area 2 has 2.052 times the odds of susceptible fish than area 1

* area 3 has 1.93 times the odds of susceptible fish than area 1

* area 4 has 0.36 times the odds of susceptible fish than area 1

* 2000 has 0.471 times the odds of susceptible fish compared to 1999. 

* 2001 has 1.579 times the odds of susceptible fish compared to 1999.

* The odds of parasite-susceptiblity changes 0.99 fold with one-unit increase in length of fish.


##### Count model coefficients

```{r warning = F}

summary(p2c)$coefficients$count %>% 
  broom::tidy() %>% 
  mutate(exp_estimate = exp(Estimate)) 

```

####Interpretation

* Adjusting for other variables, the number of parasites in area 2 is estimated to be 1.31 times the number in area 1. 

* The number of parasites in area 3 is estimated to be 1.16 times the number in area 1. 

* The number of parasites in area 4 is estimated to be 2.57 times the number in area 1.

* Adjusting for other variables, the number of parasites in year 2000 is 1.48 times the number in 1999. 

* The number of parasites in year 2001 is 0.96 times the number in 1999.

* The number of parasites changes 0.96 fold with one-unit increase in length of the fish, holding other variables constant. 
